The not so widening achievement gap: an interntational comparison of the evolution of the SES achievement gap from 2000-2015

### 2017-09-05 paper/paper.Rnw
Reorganized chunks and tested paper is running fine - still need to prettify output

<<default_conf>>=
  library(matrixStats)
  library(knitr)
  library(arm)
  library(saves)
  library(haven)
  library(PISA2000lite)
  library(PISA2003lite)
  library(PISA2006lite)
  library(PISA2009lite)
  library(PISA2012lite)
  library(intsvy)
  library(cimentadaj) # devtools::install_github("cimentadaj/cimentadaj")
  library(countrycode) # For region variable
  library(car)
  library(SAScii)
  library(inequalityintsvy) # devtools::install_github("cimentadaj/inequalityintsvy")
  library(lme4)
  library(modelr)
  library(ggrepel)
  library(xtable)
  library(ggthemes)
  library(tidyverse)

@@

Finally, the last notable pattern is that in some countries the average performance is so high that it actually overperforms the achievement gap. Only Finland, Canada and Netherlands (three very equal countries) have their average performance in standard deviations above their level of inequality, which coincidentially, are countries that have the lower leven of inequalities among the ones presented here.

<<next_steps>>=
# Next steps:

# Continue by doing the multilevel models to see what explains what. Include
# all indicators from the reardon/russian girl paper.

# Graph the increase in each country vs the increase/decrease of the economic inequality indicators
# Specially the 90/10

# Calculate how big is the gap between reading and math

# Continue with the PIRLS to see if there are specific patterns in 4th and 8th graders gap.

# Get each country trendline adjusted for the inequality indicators and place in the same country graph.

# Should I add the parent's education in the lm model to see how trends change adjusted for that?

# A weak welfare system, together with income inequality, what's their pattern?
# What if we put the school differentiation/tracking aspect in? Are there country groups based on these
# patterns.

# In countries where there is high differentiation/tracking, is there a jump in the evolution of the gap between PIRLS/TIMSS and PISA?
@

@@

\begin{figure}
\begin{center}
<<graph_all_gaps,  include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
complete_data_topbottom_80 <-
  pisa_preparer(results_math_80, results_read_80) %>%
  mutate(type = "80th/20th SES gap")

@@

######

@@

#####

complete_gaps <-
  complete_data_topbottom %>%
  bind_rows(complete_data_topbottom_80) %>%
  bind_rows(complete_data_topbottom_70)

complete_gaps %>%
  mutate(country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  filter(type_test == "math",
         country %in% countries) %>%
  ggplot(aes(as.character(wave), difference, colour = type, linetype = type, group = type)) +
  geom_line(stat = "smooth", method = "lm", size = 0.7) +
  facet_wrap(~ country, ncol = 5) +
  scale_colour_discrete(name = "Type of gaps") +
  scale_linetype_discrete(name = "Type of gaps") +
  scale_y_continuous(name = "Gaps in SD", expand = c(0, 0), lim = c(0, 3)) +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  ggthemes::theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        panel.grid.major.y = element_line(colour = "grey"),
        legend.position = "bottom") +
  ggtitle("Evolution of the different achievement gaps")
@
\caption{Evolution of the achievement gap for several gaps}
\end{center}
\end{figure}

@@

<<eval = F>>=

# 2000
# SC03Q01 School public/private
# 1 Public
# 2 Private
# 7 N/A
# 8 M/R
# 9 Mis

# 2003
# SC03Q01 (8) Public or private
# 1 Public
# 2 Private
# 7 N/A
# 8 Invalid
# 9 Miss

# 2006
# SC02Q01 (8) Public or private
#  1 Public
#  2 Private
#  7 N/A
#  8 Invalid
#  9 Missing

# 2009
# SC02Q01
# 1 public
# 2 private

# 2012
# SC01Q01 Public or private Num
# 1 Public
# 2 Private
# 7 N/A
# 8 Invalid
# 9 Missing

# 2015
# SC013Q01TA
# 1 Public
# 2 Private
# 7 N/A
# 8 Invalid
# 9 Missing
library(readxl)

tracking_data <-
  read_xlsx("./paper/data/tracking.xlsx", sheet = "all_data") %>%
  map_if(is_double, round, 2) %>%
  as_tibble

library(inequalityintsvy)

central_examination <-
  tribble(
  ~ country, ~central_examination,
  "Australia", 1, 
  "Austria", 0, 
  "Belgium", 0,
  "Bulgaria", 1,
  "Canada", rbinom(1, 1, prob = 0.51),
  "Czech Republic", 1,
  "Denmark", 1,
  "Finland", 1,
  "France", 1,
  "Germany", 0,
  "United Kingdom", 1,
  "Greece", 0,
  "Hong Kong", 1,
  "Hungary", 1,
  "Iceland", 1,
  "Ireland", 1,
  "Israel", 1,
  "Italy", 1,
  "Japan", 1,
  "Korea", 1,
  "Latvia", 1,
  "Liechtenstein", 1,
  "Luxembourg", 1,
  "Netherlands", 1,
  "New Zealand", 1,
  "Norway", 1,
  "Poland", 1,
  "Portugal", 0,
  "Russia", 1,
  "Slovakia", 1,
  "Slovenia", 1,
  "Spain", 0,
  "Sweden", 0,
  "Switzerland", 0,
  "Turkey", 1,
  "United States", 1
  )

# The tracking variables explain very nicely the difference in test scores
# for all countries pooled, up to 50% R squared.

dif_data <- function(my_data, tracking, central, which_gap = "90th/10th SES gap") {
  my_data %>%
    filter(type_test == "math", type == which_gap) %>%
    select(wave, country, difference) %>%
    rename(year = wave) %>%
    mutate(year = as.character(year)) %>%
    left_join(tracking_data, by = c("country" = "cntry_name")) %>%
    left_join(central_examination) %>%
    mutate(num_tracks = ifelse(tracks15y == 1, 1, 0) %>% as.factor,
           age_selection = selage) %>%
    filter(!is.na(num_tracks), !is.na(age_selection), !is.na(difference))

}

ready_data <- dif_data(complete_gaps, tracking_data, central_examination)

ready_data_age <-
  dif_data(complete_gaps, tracking_data, central_examination) %>%
  mutate(age_selection = ifelse(selage >= 15, 1, 0) %>% as.factor)

gaps <- c("90th/10th SES gap", "80th/20th SES gap", "70th/30th SES gap")

all_gaps_models <-
  map(gaps, function(gap) {
    ready_data_age <-
      dif_data(complete_gaps, tracking_data, central_examination, which_gap = gap) %>%
      mutate(age_selection = ifelse(selage >= 15, 1, 0) %>% as.factor)

    mod_tracking <-
      brms::brm(
        difference ~
          ztrack +
          zvoc +
          ztrack:zvoc, family = gaussian(),
      data = ready_data_age,
      warmup = 1000, iter = 2000, chains = 5
      )

    mod_tracking
})

model_data_merged <-
  map(all_gaps_models, ~ fixef(.x)[, c(1, 3, 4)] %>% as_tibble) %>%
  setNames(gaps) %>%
  enframe() %>%
  unnest(value) %>%
  mutate_if(is_double, round, 2) %>%
  transmute(model = name,
           terms = all_gaps_models[[1]] %>% fixef %>% row.names %>% rep(3),
           estimate = Estimate,
           lower = `2.5%ile`,
           upper = `97.5%ile`)

model_data_merged %>%
  ggplot(aes(terms, estimate, colour = model)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1) +
  geom_point() +
  geom_hline(yintercept = 0) +
  scale_colour_discrete(name = NULL) +
  scale_y_continuous(limits = c(-0.5, 2)) +
  ggtitle("Coef from linear model of inequality on degree of tracking and vocational tracking") +
  coord_flip()

library(brms)
# It under predicts the y value rather than over predict
# It has problems predicting the central tendency of the model
# This linear model explains the pooled 90/10 differences

mod_tracking <-
  brms::brm(
    difference ~
      ztrack +
      zvoc +
      ztrack:zvoc, family = gaussian(),
  data = ready_data_age,
  warmup = 1000, iter = 2000, chains = 5
  )

bayes_R2(mod_tracking)

plot_interaction <-
  marginal_effects(mod_tracking,
                 effects = "ztrack:zvoc",
                 int_conditions = list(zvoc = quantile))

pp_check(mod_tracking, type = "dens_overlay", nsamples = 100)
pp_check(mod_tracking, type = "scatter_avg", nsamples = 100)
pp_check(mod_tracking, type = "boxplot", nsamples = 20)
pp_check(mod_tracking, type = "error_hist", nsamples = 6)
pp_check(mod_tracking, type = "ecdf_overlay")

pp_check(mod_tracking, type = "error_scatter_avg_vs_x", x = "zvoc")
pp_check(mod_tracking, type = "error_scatter_avg_vs_x", x = "ztrack")

# I don't know how to estimate VIF's with bayesian reasoning
model_formula <- difference ~ num_tracks + age_selection + zvoc + I(length > 0)
mod_tracking_ext <-
  brms::brm(model_formula, family = gaussian(),
  data = ready_data_age,
  warmup = 1000, iter = 2000, chains = 5
  )

bayes_R2(mod_tracking_ext)

# This is the closest I could get to getting vif's:
lm(model_formula, data = ready_data_age) %>%
  vif()

new_data <-
  ready_data_age %>%
  with(crossing(country = countries,
                num_tracks = unique(num_tracks),
                age_selection = unique(age_selection),
                zvoc = mean(zvoc, na.rm = TRUE),
                length = typical(length > 0)))

predictions <- posterior_predict(mod_tracking_ext, newdata = new_data)
sd_obs <- colSds(predictions)

new_data <-
  new_data %>%
  mutate(
    pred = colMeans(predictions),
    pred_low = pred - (2 * sd_obs),
    pred_high = pred + (2 * sd_obs)
  )

sum_ready_data <-
  ready_data_age %>%
  group_by(country, num_tracks, age_selection) %>%
  summarize(avg_diff = mean(difference),
            diff_low = avg_diff - 2 * sd(difference),
            diff_high = avg_diff + 2 * sd(difference)) %>%
  ungroup()

new_data %>%
  left_join(sum_ready_data) %>%
  ggplot(aes(num_tracks %>% as.factor, avg_diff, fill = num_tracks)) +
  geom_errorbar(aes(ymin = diff_low, ymax = diff_high), width = 0.2, colour = "red") +
  geom_point(colour = "red") +
  scale_y_continuous(limits = c(0, 4)) +
  geom_col(aes(y = pred, alpha = 0.4)) +
  geom_errorbar(aes(ymin = pred_low, ymax = pred_high), width = 0.2) +
  facet_wrap(country ~ age_selection) +
  scale_fill_wsj()
@

<<>>=

# One idea I had was to make a cumulative sum of the trend,
# so the cumulative inequality over time for each country
# and see how each tracking feature explains this cumulative
# pattern, and it does so very well, with each one
# separately (because I have only 30 observations) explaining
# about 22% each. Age selection, up to 30% alone.

vars_unique <- c("length", "num_tracks", "age_selection", "central_examination", "ztrack",
                 "zvoc")

ready_data_two <-
  ready_data %>%
  group_by(country) %>%
  summarize_at(vars(vars_unique), unique) %>%
  mutate(cum_diff = ready_data %>% group_by(country) %>% summarize(m = sum(difference)) %>% pull(m))

mod_tracking <-
  ready_data_two %>%
  lm(cum_diff ~ ztrack + zvoc, data = .) %>%
  arm::display(detail = T)

mod_tracking <-
  ready_data_two %>%
  filter(!is.na(age_selection)) %>%
  brm(
    cum_diff ~ I(age_selection >= 15),
    family = gaussian(),
    data = .,
    warmup = 1000, iter = 2000, chains = 5
  )

bayes_R2(mod_tracking)

pp_check(mod_tracking, type = "dens_overlay", nsamples = 100)
pp_check(mod_tracking, type = "ecdf_overlay", nsamples = 10)
pp_check(mod_tracking, type = "boxplot", nsamples = 30)
pp_check(mod_tracking, type = "stat", nsamples = 200) # very nice to compare central tendency
pp_check(mod_tracking, type = "error_scatter_avg", nsamples = 1000)

pp_check(mod_tracking, type = "loo_pit")
# over prediction
# central tendency is not captured very well
# more under-prediction than over prediction

@@

<<next_steps>>=
# Next steps:

# Continue by doing the multilevel models to see what explains what. Include
# all indicators from the reardon/russian girl paper.

# Graph the increase in each country vs the increase/decrease of the economic inequality indicators
# Specially the 90/10

# Calculate how big is the gap between reading and math

# Continue with the PIRLS to see if there are specific patterns in 4th and 8th graders gap.

# Get each country trendline adjusted for the inequality indicators and place in the same country graph.

# Should I add the parent's education in the lm model to see how trends change adjusted for that?

# A weak welfare system, together with income inequality, what's their pattern?
# What if we put the school differentiation/tracking aspect in? Are there country groups based on these
# patterns.

# In countries where there is high differentiation/tracking, is there a jump in the evolution of the gap between PIRLS/TIMSS and PISA?
@

>>>>

<<default_conf>>=
  library(matrixStats)
  library(knitr)
  library(arm)
  library(saves)
  library(haven)
  library(PISA2000lite)
  library(PISA2003lite)
  library(PISA2006lite)
  library(PISA2009lite)
  library(PISA2012lite)
  library(intsvy)
  library(cimentadaj) # devtools::install_github("cimentadaj/cimentadaj")
  library(countrycode) # For region variable
  library(car)
  library(SAScii)
  library(inequalityintsvy) # devtools::install_github("cimentadaj/inequalityintsvy")
  library(lme4)
  library(modelr)
  library(ggrepel)
  library(xtable)
  library(ggthemes)
  library(readxl)
  library(inequalityintsvy)
  library(brms)
  library(tidyverse)

@@

Finally, the last notable pattern is that in some countries the average performance is so high that it actually overperforms the achievement gap. Only Finland, Canada and Netherlands (three very equal countries) have their average performance in standard deviations above their level of inequality, which coincidentially, are countries that have the lower level of inequalities among the ones presented here.

% 90/80/70 gaps merged

@@

<<merge_all_gaps>>=
complete_data_topbottom_80 <-
  pisa_preparer(results_math_80, results_read_80) %>%
  mutate(type = "80th/20th SES gap")

@@

####

@@

####

complete_gaps <-
  complete_data_topbottom %>%
  bind_rows(complete_data_topbottom_80) %>%
  bind_rows(complete_data_topbottom_70)
@

% end here

% <<>>=
% # 2000
% # SC03Q01 School public/private
% # 1 Public
% # 2 Private
% # 7 N/A
% # 8 M/R
% # 9 Mis
% 
% # 2003
% # SC03Q01 (8) Public or private
% # 1 Public
% # 2 Private
% # 7 N/A
% # 8 Invalid
% # 9 Miss
% 
% # 2006
% # SC02Q01 (8) Public or private
% #  1 Public
% #  2 Private
% #  7 N/A
% #  8 Invalid
% #  9 Missing 
% 
% # 2009
% # SC02Q01
% # 1 public
% # 2 private
% 
% # 2012
% # SC01Q01 Public or private Num
% # 1 Public
% # 2 Private
% # 7 N/A
% # 8 Invalid
% # 9 Missing
% 
% # 2015
% # SC013Q01TA
% # 1 Public
% # 2 Private
% # 7 N/A
% # 8 Invalid
% # 9 Missing
% 
% # central_examination <-
% #   tribble(
% #   ~ country, ~central_examination,
% #   "Australia", 1, 
% #   "Austria", 0, 
% #   "Belgium", 0,
% #   "Bulgaria", 1,
% #   "Canada", rbinom(1, 1, prob = 0.51),
% #   "Czech Republic", 1,
% #   "Denmark", 1,
% #   "Finland", 1,
% #   "France", 1,
% #   "Germany", 0,
% #   "United Kingdom", 1,
% #   "Greece", 0,
% #   "Hong Kong", 1,
% #   "Hungary", 1,
% #   "Iceland", 1,
% #   "Ireland", 1,
% #   "Israel", 1,
% #   "Italy", 1,
% #   "Japan", 1,
% #   "Korea", 1,
% #   "Latvia", 1,
% #   "Liechtenstein", 1,
% #   "Luxembourg", 1,
% #   "Netherlands", 1,
% #   "New Zealand", 1,
% #   "Norway", 1,
% #   "Poland", 1,
% #   "Portugal", 0,
% #   "Russia", 1,
% #   "Slovakia", 1,
% #   "Slovenia", 1,
% #   "Spain", 0,
% #   "Sweden", 0,
% #   "Switzerland", 0,
% #   "Turkey", 1,
% #   "United States", 1
% #   )
% 
% @
% 
% <<complete_tracking_model>>=
% tracking_data <-
%   read_xlsx("./data/tracking.xlsx", sheet = "all_data") %>%
%   map_if(is_double, round, 2) %>%
%   as_tibble
% 
% dif_data <- function(my_data, tracking, central, which_gap = "90th/10th SES gap") {
%   my_data %>%
%     filter(type_test == "math", type == which_gap) %>%
%     select(wave, country, difference) %>%
%     rename(year = wave) %>%
%     mutate(year = as.character(year)) %>%
%     left_join(tracking_data, by = c("country" = "cntry_name")) %>%
%     left_join(central_examination) %>%
%     mutate(num_tracks = ifelse(tracks15y == 1, 1, 0) %>% as.factor,
%            age_selection = selage) %>%
%     filter(!is.na(num_tracks), !is.na(age_selection), !is.na(difference))
% 
% }
% formula_sequence <- function(dv, ivs, random_effect) {
%   formula_seq <-
%     lapply(1:length(ivs), function(x) {
%       paste0(dv ," ~ ", paste0(ivs[1:x], collapse = " + ") %>% paste0(" + ", random_effect))
%     }) %>%
%     lapply(as.formula)
% 
%   formula_seq  
% }
% stan_model_builder <- function(dv, iv, random, data) {
%   all_formulas <- formula_sequence(dv, iv, random)
% 
%   mod_tracking <-
%     map(all_formulas, ~ {
%       brms::brm(
%         .x,
%         family = gaussian(),
%         data = data,
%         warmup = 1000, iter = 2000, chains = 5
%       )
%     })
%   
% }
% stan_extractor <- function(models) {
%   coef_list <-
%     map(models, ~ {
%     broom::tidy(.x) %>%
%       .[grepl("^b_", .$term), ] %>%
%       map_if(is_double, round, 2) %>%
%       as_tibble %>%
%       transmute(term,
%                 estimate = paste0(estimate, " (", lower, "/", upper, ")"))
%   })
%   
%   coef_list
% }
% stan_table_builder <- function(models_extracted) {
%   intermediate_table <-
%     reduce(models_extracted, full_join, by = "term") %>%
%     setNames(c(" ", paste("Model", seq_len(models_extracted %>% length))))
%   
%   row_order <-
%     setdiff(seq_len(nrow(intermediate_table)),
%             grep("Intercept", intermediate_table$` `)) %>%
%     `c`(grep("Intercept", intermediate_table$` `))
% 
%     intermediate_table[row_order, ]
% }
% stan_table <- function(models) {
% 
%   table_ready <-
%     models %>%
%     stan_extractor() %>%
%     stan_table_builder()
%   
%   table_ready
% }
% 
% table_details <- function(models) {
%   
%   random_effect <- nrow(models[[length(models)]]$ranef) > 0
%   
%   where_to_place <- 
%     models[[length(models)]]$formula$formula %>%
%     as.character %>%
%     .[-(1:2)] %>% 
%     stringr::str_split(" \\+ ") %>%
%     pluck(1) %>%
%     .[if (random_effect) -length(.) else 1:length(.)] %>%
%     length
%   
%   r_square <-
%     paste0("\\hline ",
%            paste0("R square: & ",
%                   paste0(map(models, ~ bayes_R2(.x)[1] %>% round(2) * 100) %>% paste0("\\%"),
%                                      collapse = " & ")), " \\\\")
%   n_obs <-
%     paste0(paste0(" Sample size: & ",
%                   paste0(map(models, nobs), collapse = " & ")), " \\\\ \\hline")
%   
%   command_to_row <- paste0(r_square, n_obs)
%   addtorow_two <- list(pos = list(pos = where_to_place + 1), command = command_to_row)
%   
%   addtorow_two
% }
% 
% ready_data_age <-
%   dif_data(complete_gaps, tracking_data, central_examination) %>%
%   mutate(age_selection = ifelse(selage >= 15, 1, 0) %>% as.factor)
% 
% all_track_models <-
%   stan_model_builder(
%   dv = "difference",
%   iv = c("num_tracks", "age_selection", "I(length > 0)", "zvoc"),
%   random = "(1 | year)",
%   data = ready_data_age
% )
% 
% all_track_models %>%
%   stan_table() %>%
%   mutate(" " = c(
%     "Only 1 track",
%     "Age selection >= 15",
%     "% of curric tracked",
%     "Vocational Index",
%     "Intercept"
%   )) %>%
%   xtable() %>%
%   print.xtable(
%   include.rownames = FALSE,
%   add.to.row = table_details(all_track_models)
%   )
% @
% 
% <<graph_complete_tracking_model>>=
% # This is the closest I could get to getting vif's fro the last model:
% 
% lm(difference ~ num_tracks + age_selection + I(length > 0) + zvoc,
%    data = ready_data_age) %>%
%   vif()
% 
% mod_tracking_ext <-
%   brms::brm(formula_seq[[4]],
%             family = gaussian(),
%             data = ready_data_age,
%             warmup = 1000, iter = 2000, chains = 5,
%             control = list(adapt_delta = 0.90)
%   )
% 
% new_data <-
%   ready_data_age %>%
%   with(crossing(country = countries,
%                 num_tracks = unique(num_tracks),
%                 age_selection = unique(age_selection),
%                 zvoc = mean(zvoc, na.rm = TRUE),
%                 length = typical(length > 0),
%                 year = "2009"))
% 
% predictions <- posterior_predict(mod_tracking_ext, newdata = new_data)
% sd_obs <- colSds(predictions)
% 
% new_data <-
%   new_data %>%
%   mutate(
%     pred = colMeans(predictions),
%     pred_low = pred - (2 * sd_obs),
%     pred_high = pred + (2 * sd_obs)
%   )
% 
% sum_ready_data <-
%   ready_data_age %>%
%   group_by(country, num_tracks, age_selection) %>%
%   summarize(avg_diff = mean(difference),
%             diff_low = avg_diff - 2 * sd(difference),
%             diff_high = avg_diff + 2 * sd(difference)) %>%
%   ungroup()
% 
% new_data %>%
%   left_join(sum_ready_data) %>%
%   mutate(age_selection = ifelse(age_selection == 1, ">= 15", "< 15")) %>%
%   ggplot(aes(num_tracks, avg_diff, fill = num_tracks)) +
%   geom_errorbar(aes(ymin = diff_low, ymax = diff_high), width = 0.2, colour = "red") +
%   geom_point(colour = "red") +
%   geom_col(aes(y = pred, alpha = 0.4)) +
%   geom_errorbar(aes(ymin = pred_low, ymax = pred_high), width = 0.2) +
%   facet_wrap(country ~ age_selection) +
%   scale_y_continuous(name = "90/10 SD gap", limits = c(0, 4)) +
%   scale_alpha(guide = F) +
%   scale_x_discrete(name = "Number of tracks", labels = c("> 1 track", "1 track")) +
%   scale_fill_manual(values = c("orangered2", "blue3"), guide = F) +
%   ggtitle("Predicted 90/10 gap for different combinations of number of tracks and age of selection")
% @
% 
% <<counterfactual_tracking>>=
% pred_avg_diff<-
%   new_data %>%
%   left_join(sum_ready_data) %>%
%   filter(!is.na(avg_diff) | num_tracks == 1 & age_selection == 1) %>%
%   select(country, num_tracks, age_selection, pred, avg_diff)
% 
% track_should_change <- pred_avg_diff$country[pred_avg_diff$country %>% duplicated]
% 
% avg_diff_data <-
%   pred_avg_diff %>%
%   filter(country %in% track_should_change) %>%
%   transmute(country,
%             avg_diff = ifelse(avg_diff %>% is.na, pred, avg_diff),
%             combn = paste0(num_tracks, "_", age_selection)) %>%
%   filter(combn != "1_1")
% 
% final_avg_diff_table <-
%   avg_diff_data %>%
%   left_join(pred_avg_diff %>%
%               filter(country %in% track_should_change & (num_tracks == 1 & age_selection == 1)) %>%
%               select(country, pred), by = "country") %>%
%   select(country, combn, avg_diff, pred) %>%
%   separate(combn, c("num_tracks", "age_selection"), sep = "_") %>%
%   transmute(
%     Country = country,
%     `Number of tracks` = ifelse(num_tracks == 0, "More than one track", "One track"),
%     `Age of selection` = ifelse(age_selection == 1, "15 or more", "Less than 15"),
%     `Inequality` = avg_diff %>% round(3),
%     `Predicted` = pred %>% round(3),
%     `% reduction` = (((`Inequality` - `Predicted`)/ `Inequality`) * 100) %>%
%       ceiling %>% paste0("%")
%     )
% 
% xtable(final_avg_diff_table) %>% print.xtable(
%   include.rownames = FALSE
%   )
% 
% @
% 
% <<interaction_model>>=
% ready_data <- dif_data(complete_gaps, tracking_data, central_examination)
% 
% stan_models <-
%   stan_model_builder(
%   dv = "difference",
%   iv = c("ztrack", "zvoc", "ztrack:zvoc"),
%   random = "(1 | year)",
%   data = ready_data
%   )
% 
% table_ready <-
%   stan_models %>%
%   stan_table() %>%
%   mutate(" " = c(
%     "Tracking Index",
%     "Vocational Index",
%     "Trackin * Vocational Index",
%     "Intercept"
%   )) %>%
%   xtable() %>%
%   print.xtable(
%     include.rownames = FALSE,
%     add.to.row = table_details(stan_models)
%   )
% 
% plot_interaction <-
%   marginal_effects(stan_models[[length(stan_models)]],
%                  effects = "ztrack:zvoc",
%                  int_conditions = list(zvoc = quantile))
% 
% plot(plot_interaction, plot = FALSE)[[1]] +
%   labs(x = "Tracking Index (SD)", y = "Difference in test scores (SD)") +
%   scale_colour_discrete(
%     name = "Vocational index \n quantiles",
%     labels = seq(100, 0, -25) %>% paste0("%")
%     ) +
%   scale_fill_discrete(guide = FALSE) +
%   ggtitle("Interaction between Tracking and Vocational index")
% 
% # pp_check(mod_tracking, type = "dens_overlay", nsamples = 100)
% # pp_check(mod_tracking, type = "scatter_avg", nsamples = 100)
% # pp_check(mod_tracking, type = "boxplot", nsamples = 20)
% # pp_check(mod_tracking, type = "error_hist", nsamples = 6)
% # pp_check(mod_tracking, type = "ecdf_overlay")
% # 
% # pp_check(mod_tracking, type = "error_scatter_avg_vs_x", x = "zvoc")
% # pp_check(mod_tracking, type = "error_scatter_avg_vs_x", x = "ztrack")
% @
% 
% <<all_gaps_plot>>=
% gaps <- c("90th/10th SES gap", "80th/20th SES gap", "70th/30th SES gap")
% all_gaps_models <-
%   map(gaps, function(gap) {
%     ready_data_age <-
%       dif_data(complete_gaps, tracking_data, central_examination, which_gap = gap) %>%
%       mutate(age_selection = ifelse(selage >= 15, 1, 0) %>% as.factor)
%     
%     mod_tracking <-
%       brms::brm(
%         difference ~
%           ztrack +
%           zvoc +
%           ztrack:zvoc +
%           (1|year), family = gaussian(),
%       data = ready_data_age,
%       warmup = 1000, iter = 2000, chains = 5,
%       control = list(adapt_delta = 0.95)
%       )
%     
%     mod_tracking
% })
% 
% model_data_merged <-
%   map(all_gaps_models, ~ fixef(.x)[, c(1, 3, 4)] %>% as_tibble) %>%
%   setNames(gaps) %>%
%   enframe() %>%
%   unnest(value) %>%
%   mutate_if(is_double, round, 2) %>%
%   transmute(model = name,
%            terms = all_gaps_models[[1]] %>% fixef %>% row.names %>% rep(3),
%            estimate = Estimate,
%            lower = `2.5%ile`,
%            upper = `97.5%ile`)
% 
% model_data_merged %>%
%   ggplot(aes(terms, estimate, colour = model)) +
%   geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1) +
%   geom_point() +
%   geom_hline(yintercept = 0) +
%   scale_colour_discrete(name = NULL) +
%   scale_y_continuous(limits = c(-0.5, 2)) +
%   scale_x_discrete(labels = c("Intercept",
%                               "Tracking Index",
%                               "Vocational * Tracking Index",
%                               "Vocational Index")) +
%   labs(y = "Estimate from linear model", x = "Predictors") +
%   ggtitle("Estimation of tracking/vocational models for different gaps") +
%   coord_flip()
% 
% # It under predicts the y value rather than over predict
% # It has problems predicting the central tendency of the model
% # This linear model explains the pooled 90/10 differences
% @
% 
% <<cumulative_modeling_table>>=
% 
% # One idea I had was to make a cumulative sum of the trend,
% # so the cumulative inequality over time for each country
% # and see how each tracking feature explains this cumulative
% # pattern, and it does so very well, with each one
% # separately (because I have only 30 observations) explaining
% # about 22% each. Age selection, up to 30% alone.
% 
% vars_unique <- c("length", "num_tracks", "age_selection", "central_examination", "ztrack",
%                  "zvoc")
% 
% all_gaps_models_cum <-
%   map(gaps, function(gap) {
%     
%     gap_data <- dif_data(complete_gaps, tracking_data, central_examination, which_gap = gap)
%     
%     ready_data_age <-
%       gap_data %>%
%       mutate(age_selection = ifelse(selage >= 15, 1, 0) %>% as.factor) %>%
%       group_by(country) %>%
%       summarize_at(vars(vars_unique), unique) %>%
%       mutate(cum_diff =
%                gap_data %>%
%                group_by(country) %>%
%                summarize(m = sum(difference)) %>%
%                pull(m))
%     
%     mod_tracking <-
%       ready_data_age %>%
%       filter(!is.na(age_selection)) %>%
%       brm(
%         cum_diff ~ ztrack + zvoc,
%         family = gaussian(),
%         data = .,
%         warmup = 1000, iter = 2000, chains = 5
%       )
% 
%     mod_tracking
% })
% 
% all_gaps_models_cum %>%
%   stan_table() %>%
%   mutate(" " = c(
%     "Track Index",
%     "Vocational Index",
%     "Intercept"
%   )) %>%
%   setNames(c(" ", gsub(" SES gap", "", gaps))) %>%
%   xtable() %>%
%   print.xtable(
%     include.rownames = FALSE,
%     add.to.row = table_details(all_gaps_models_cum)
%   )
% 
% @
% 
% <<next_steps>>=
% # Next steps:
% 
% # Continue by doing the multilevel models to see what explains what. Include
% # all indicators from the reardon/russian girl paper.
%   
% # Graph the increase in each country vs the increase/decrease of the economic inequality indicators
% # Specially the 90/10
% 
% # Calculate how big is the gap between reading and math
%   
% # Continue with the PIRLS to see if there are specific patterns in 4th and 8th graders gap.
%   
% # Get each country trendline adjusted for the inequality indicators and place in the same country graph.
% 
% # Should I add the parent's education in the lm model to see how trends change adjusted for that?
% 
% # A weak welfare system, together with income inequality, what's their pattern?
% # What if we put the school differentiation/tracking aspect in? Are there country groups based on these
% # patterns.
% 
% # In countries where there is high differentiation/tracking, is there a jump in the evolution of the gap between PIRLS/TIMSS and PISA?
% @

\begin{figure}
\begin{center}
<<graph_all_gaps, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
complete_gaps %>%
  mutate(country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  filter(type_test == "math",
         country %in% countries) %>%
  ggplot(aes(as.character(wave), difference, colour = type, linetype = type, group = type)) +
  geom_line(stat = "smooth", method = "lm", size = 0.7) +
  facet_wrap(~ country, ncol = 5) +
  scale_colour_discrete(name = "Type of gaps") +
  scale_linetype_discrete(name = "Type of gaps") +
  scale_y_continuous(name = "Gaps in SD", expand = c(0, 0), lim = c(0, 3)) +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  ggthemes::theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        panel.grid.major.y = element_line(colour = "grey"),
        legend.position = "bottom") +
  ggtitle("Evolution of the different achievement gaps")
@
\caption{Evolution of the achievement gap for several gaps}
\end{center}
\end{figure}

@@

<<eval = F>>=
# pp_check(mod_tracking, type = "dens_overlay", nsamples = 100)
# pp_check(mod_tracking, type = "ecdf_overlay", nsamples = 10)
# pp_check(mod_tracking, type = "boxplot", nsamples = 30)
# pp_check(mod_tracking, type = "stat", nsamples = 200) # very nice to compare central tendency
# pp_check(mod_tracking, type = "error_scatter_avg", nsamples = 1000)
# 
# pp_check(mod_tracking, type = "loo_pit")
# over prediction
# central tendency is not captured very well
# more under-prediction than over prediction

@@
